---
title: "Team Stats"
---

```{r}
#| echo: false
#| message: false
library(dplyr)
library(knitr)
library(plotly)

ordinal <- function(n) {
  if (is.na(n)) return("—")
  suffix <- if (n %% 100 %in% 11:13) "th" else
    switch(as.character(n %% 10), "1" = "st", "2" = "nd", "3" = "rd", "th")
  paste0(n, suffix)
}

stats_path <- "data/team-stats.csv"
events_path <- "data/team-event-results.csv"
have_data  <- file.exists(stats_path)
stats  <- if (have_data) read.csv(stats_path) else data.frame()
events <- if (file.exists(events_path)) read.csv(events_path) else data.frame()
```

::: {.callout-note}
**About this data.** These numbers come from [Duosmium](https://github.com/Duosmium/duosmium), a community-maintained, open-source archive of Science Olympiad results — not an official database. Two things follow from that:

- **Coverage is incomplete.** It only includes tournaments someone
  chose to upload, so a season or an invitational can be missing
  entirely (see `R/update_duosmium_stats.R` for which ones we
  currently pull from).
- **Individual results can be wrong.** Uploads occasionally have
  transcription errors that never got corrected. If a number here
  doesn't match what you remember from awards ceremony, trust your
  own memory — and feel free to flag it so we can note the
  discrepancy or exclude that file.

Best-invitational placement isn't shown as a headline stat, since
invitational field sizes and formats vary too much to compare fairly
across tournaments — invitational results still count toward the
event medal total below.
:::

```{r}
#| echo: false
#| results: asis
if (!have_data || nrow(stats) == 0) {
  cat("_No data yet — run `Rscript R/update_duosmium_stats.R` to pull results from Duosmium._")
}
```

::: {.grid}

::: {.g-col-6 .g-col-md-3 .stat-card}
```{r}
#| echo: false
#| results: asis
best_reg <- if (have_data && any(stats$level == "Regionals", na.rm = TRUE)) {
  min(stats$rank[stats$level == "Regionals"], na.rm = TRUE)
} else NA
cat("<div class='stat-number'>", ordinal(best_reg), "</div>")
```
<div class="stat-label">Best Regional Finish</div>
:::

::: {.g-col-6 .g-col-md-3 .stat-card}
```{r}
#| echo: false
#| results: asis
best_state <- if (have_data && any(stats$level == "States", na.rm = TRUE)) {
  min(stats$rank[stats$level == "States"], na.rm = TRUE)
} else NA
cat("<div class='stat-number'>", ordinal(best_state), "</div>")
```
<div class="stat-label">Best State Finish</div>
:::

::: {.g-col-6 .g-col-md-3 .stat-card}
```{r}
#| echo: false
#| results: asis
total_medals <- if (have_data) sum(stats$event_medals, na.rm = TRUE) else 0
cat("<div class='stat-number'>", total_medals, "</div>")
```
<div class="stat-label">Career Event Medals</div>
:::

::: {.g-col-6 .g-col-md-3 .stat-card}
```{r}
#| echo: false
#| results: asis
n_tournaments <- if (have_data) nrow(stats) else 0
cat("<div class='stat-number'>", n_tournaments, "</div>")
```
<div class="stat-label">Tournaments on Record</div>
:::

:::

## Our Best Events

```{r}
#| echo: false
#| warning: false
if (nrow(events) == 0) {
  cat("Run the update script to populate this chart.")
} else {
  event_summary <- events |>
    group_by(event) |>
    summarise(
      medals      = sum(medal, na.rm = TRUE),
      best_place  = min(place, na.rm = TRUE),
      avg_place   = round(mean(place, na.rm = TRUE), 1),
      times       = n(),
      .groups = "drop"
    ) |>
    arrange(desc(medals), best_place) |>
    slice_head(n = 15)

  plot_ly(
    event_summary,
    x = ~reorder(event, medals),
    y = ~medals,
    type = "bar",
    orientation = "v",
    marker = list(color = "#E8720C"),
    text = ~paste0(
      event, "<br>Medals: ", medals,
      "<br>Best finish: ", best_place,
      "<br>Avg finish: ", avg_place,
      "<br>Competed: ", times, "x"
    ),
    hoverinfo = "text"
  ) |>
    layout(
      xaxis = list(title = "", tickangle = -45),
      yaxis = list(title = "Career medals", tick0 = 0, dtick = 1),
      margin = list(b = 140)
    )
}
```

## Tournament History

```{r}
#| echo: false
if (have_data && nrow(stats) > 0) {
  stats |>
    arrange(desc(year)) |>
    transmute(
      Year = year,
      Tournament = tournament,
      Level = level,
      Div = division,
      Rank = paste0(rank, " / ", n_teams),
      Trophy = ifelse(trophy, "🏆", ""),
      `Event Medals` = event_medals
    ) |>
    kable()
} else {
  cat("Run the update script to populate this table.")
}
```